---
title: "25 years of Social Work Research"
author: "Summary of Data Cleaning Procedures"
output: pdf_document
---

# Overview

This document summarizes data cleaning procedures of the data file processed by the `ebscoBWR` fucntion.  It should be noted that that `ebscoBWR` function merged the search results from the first- and second-level searches performed on the three databases (psycInfo, Social Science Abstracts, and Social Work Abstract) hosted by the EbscoPlatform.  This function merged all the search results, transformed the data into a structured data file, and removed all duplicate article records.  No further transformations or cleaning was performed by the `ebscoBWR` function.  Provided below are basic descriptive summaries of the results from the full search before further cleaning.  

The main file for analysis is `fullDF.R`, and a separate file is just for psycInfo `psycInfoDF.R`.  The same cleaning procedures need to be performed on both data files prior to analysis.  


# Initialization of workspace

```{r initialization_dataprocessing, warning=FALSE, message=FALSE, eval=TRUE}
# Clear workspace
rm(list=ls())   

# Load required packages
library(ggplot2)
library(dplyr)

# Read data files

#load("/Users/beperron/Git/SocialWorkResearch/Data/ebscoFULL.R")
#load("/Users/beperron/Git/SocialWorkResearch/Data/proQuest.R")
#load("/Users/beperron/Git/SocialWorkResearch/Data/ebscoPsycInfo.R")

load("/Users/SSW/Documents/GitHub/SocialWork100Years/Data/ebscoFULL.R")
load("/Users/SSW/Documents/GitHub/SocialWork100Years/Data/proQuest.R")
load("/Users/SSW/Documents/GitHub/SocialWork100Years/Data/ebscoPsycInfo.R")

```

Everything written after this point was for the ebsco full data file (psycinof, swa, ssa).  The R objects have been all re-written and renamed.  The first thing to do is to append the ebscoFULL and proQuest objects.  Here is my recommended steps:

1.  Create a variable that takes the maximum values of the articleID from ebscoFULL
2.  Add the max value to the article ID of proQuest
3.  Write a function that matches the titles of ebscoFull with proQuest, grabs the ID, and then, eliminates the duplicates.  We have all the matching code in the original functions that you can grab.  
4.  Consider drawing on the pmatch.R function in the BWR folder.  This function forces all to lower, strips all the white space, and takes a subset of the characters (e.g., ~25 I think).  The advantage of this is to avoid the incorrect matching when colons are used.  Same title, but you get the idea of where things can go awry with title matching.  Test it extensively -- e.g., alphebatizing is also helpful.  

Social work research can be really:  Bad
Social work research can be really: Bad
Social Work research can be really: bad
Social work research can be really: Bad
Social work research can be really: bad
Social Work research can be really:  bad

5.  Do necessary search and replace of subsequent code to get it working with the fully appended data file.

#Combines EbscoHost and ProQUest wrangles and eliminates duplicate article titles

```{r}

library(dplyr)

  #Creates variable with highest article ID from ebscohost database
  ebscoIDmax<-max(ebscoBWR.df$articleID)

  #Recodes article ID in ProQuest database in advance of merger with ebscohost
  proQuestBWR.df<-mutate(proQuestBWR.df, articleID=articleID+ebscoIDmax)

  #Creates data frames that contain only article ID and titles from respective
  #databases
  ebscoTitles<-filter(ebscoBWR.df, attributes=="article")
  proTitles<-filter(proQuestBWR.df, attributes=="article")

  #Combines titles into single data frame
  jointTitles<-rbind(ebscoTitles, proTitles)

  #Forces article title text to lower case
  jointTitles$record<-tolower(jointTitles$record)
  
  #Strips out white space and punctuation to ensure small differences
  #in data entry are detected during duplicate elimination
  jointTitles$record<-gsub(" ","", jointTitles$record)
  
  #punct <- '[]\\?!\"\'#$%&(){}+*/:;,._`|~\\[<=>@\\^-]'
  #JointTitles$record<-gsub(punct, "", JointTitles$record)  
  jointTitles$record<-gsub("[[:punct:]]", "", jointTitles$record)

  jointTitles<- jointTitles[duplicated(jointTitles$record), ]

  #Screen out duplicated records by articleID. The articleID must be used
  #because the duplicate title contains other article attributes
  jointTitles.dup.ID <- jointTitles$articleID
  
  #Combines EbscoHost and ProQuest Wrangles, and strips out article attributes 
  #from duplicate entries
  combined.DF<-rbind(ebscoBWR.df, proQuestBWR.df)
  full.df<-combined.DF[!(combined.DF$articleID %in% jointTitles.dup.ID), ]  



```





# Set the working directory
#setwd("/Users/beperron/Git/SocialWorkResearch/Data")
#setwd("/Users/SSW/Documents/GitHub/SocialWork100Years/Data")
```

# Pre-processing summary
```{r echo=FALSE}
full.df$record <- ifelse(full.df$record == 
        "Social Work & Social Sciences Review",     # Spelling difference
        "Social Work and Social Sciences Review", 
                        full.df$record)

full.df$record <- ifelse(full.df$record == 
        "Social Work with Groups",                  # Spelling difference
        "Social Work With Groups", 
                        full.df$record)

full.df$record <- ifelse(full.df$record == 
        "Journal of Sociology & Social Welfare",    # Spelling difference
        "Journal of Sociology and Social Welfare",  
                        full.df$record)

full.df$record <- ifelse(full.df$record == 
    "The British Journal of Social Work",          # Spelling difference
        "British Journal of Social Work",           
                        full.df$record)

full.df$record <- ifelse(full.df$record ==
    "Revista de Asistenta Sociala (Social Work Review)",
        "Social Work Review",
                        full.df$record)

full.df$record <- gsub("Arete", "xxx", full.df$record)
full.df$record <- gsub("Aret.{1,3}", "Arete", full.df$record)
full.df$record <- gsub("xxx", "Arete", full.df$record) 
                                      #Replaces special character       
                                      #in Arete and preserves other formatting
      
full.df$record <- ifelse(full.df$record ==
    "Journal of Gay and Lesbian Social Services",     #Spelling difference
        "Journal of Gay & Lesbian Social Services",
                        full.df$record)

full.df$record <- ifelse(full.df$record ==
    "Canadian Social Work Review/Revue canadienne de service social",
        "Canadian Social Work Review",
                        full.df$record)

full.df$record <- ifelse(full.df$record ==
    "INDIAN JOURNAL OF SOCIAL WORK",                    #Spelling difference
        "Indian Journal of Social Work",
                      full.df$record)

full.df$record <- ifelse(full.df$record ==
    "The Indian Journal of Social Work",              #Spelling difference
        "Indian Journal of Social Work",
                            full.df$record)

full.df$record <- ifelse(full.df$record ==
    "Asia Pacific Journal of Social Work",              #Spelling difference
        "Asia Pacific Journal of Social Work and Development",
                            full.df$record)

full.df$record <- ifelse(full.df$record ==
    "Social Work & Society",                           #Spelling difference
        "Social Work and Society",
                            full.df$record)

full.df$record <- ifelse(full.df$record ==
    "Journal of Social Work in Disability & Rehabilitation",
        "Journal of Social Work in Disability and Rehabilitation", 
                            full.df$record)           #Spelling difference

full.df$record <- ifelse(full.df$record ==
    "Hong Kong Journal of Social Work",
        "The Hong Kong Journal of Social Work", 
                            full.df$record)           #Spelling difference

full.df$record <- ifelse(full.df$record ==
    "The Journal of Baccalaureate Social Work",
        "Journal of Baccalaureate Social Work", 
                            full.df$record)           #Spelling difference

full.df$record <- ifelse(full.df$record ==
    "Journal of Practice Teaching in Health and Social Work",
        "Journal of Practice Teaching in Social Work and Health", 
                            full.df$record)           #Spelling difference

full.df$record <- ifelse(full.df$record ==
    "Professional Development: The International Journal of Continuing Social Work Education",
        "Professional Development", 
                            full.df$record)           #Spelling difference

```


Provided below is a summary of the datafile processed by `ebscoBWR` function before any cleaning procedures specifically for this project were employed, with the exception of resolving spelling errors or differences.  For example, _Social Work With Groups_ was exported as a separate journal than _Social Work with Groups_ (note the capitalization difference in the word _with_.)

```{r preprocessing, comment=NA, echo=TRUE} 
number.of.articles <- length(which(full.df$attributes == "article"))
range.of.years <- range(full.df$record[full.df$attributes == "pubYear"])
```


```{r}
# Number of articles
cat(number.of.articles)
```

```{r}
# Range of years
cat(range.of.years)
```


```{r date, echo=FALSE, comment=NA}
unique.titles <- filter(full.df, attributes == "journal") %>%
    group_by(record) %>%
    summarise(
        N = n())
print.data.frame(unique.titles, rowNames=FALSE)
```

# Data cleaning

Inspection of the data indicated notable delays in indexing for the years 2014 and 2015.  Therefore, these years were excluded from the current study.  Thus, this study includes journals indexed through the end of 2013.  

## Year specification

```{r}

year.25 <- as.character(c(1900:1987, 2014, 2015))
years <- c("2014", "2015")
#ebsco.remove <- full.df[full.df$record %in% years, ]
ebsco.remove <- full.df[full.df$record %in% year.25, ]
ebsco.remove <- ebsco.remove[, "articleID"]
full.df <- full.df[!(full.df$articleID %in% ebsco.remove),]
```

## Exclusion of journals not meeting supplemental search terms

The search for journals involved a two-stage process.  It began by extracting all records from journals based on the social work journal list of Hodge and L (APA cite).  Journals were excluded if they did not contain one of the major search terms from the supplemental list.  The majority of these journals were captured in the supplemental search query because the journal used one of the supplemental search terms in a special issue.  The mission and scope of journals were also reviewed when it was unclear whether the journal was specific to social work research (e.g., _The Clinical Supervisor_).  The supplemental search terms included:

+ social work
+ social welfare
+ social casework (or social case work)
+ social service
+ human service
+ social development 
+ social environment


```{r exclusion1, eval=TRUE}
exclusions.1 <- c(
    "American Journal of Preventive Medicine", 
    "Behavior Modification", 
    "Brain and Cognition",
    "Canadian Journal of Community Mental Health",
    "Canadian Journal on Aging",
    "Early Child Development and Care",
    "Early Education and Development",
    "Employee Assistance Quarterly",
    "Evaluation and Program Planning",
    "General Hospital Psychiatry",
    "International Review of Social Research",
    "Journal of Applied Behavioral Science",
    "Journal of Applied Rehabilitation Counseling",
    "Journal of Community & Applied Social Psychology",
    "Journal of Health & Human Services Administration",
    "Journal of Nonverbal Behavior",
    "Journal of Psychosomatic Research",
    "Learning Disabilities Research & Practice",
    "Professional Development in Education",
    "PROFILE Issues in Teachers' Professional Development",
    "Psychotherapy",
    "Systems Research and Behavioral Science",
    "Teaching and Teacher Education",
    "The Scientific Review of Mental Health Practice",
    "The Clinical Supervisor",
    "Arete: Revista de Filosofia",
    "Arete Revista de Filosofia")

ebsco.remove <- full.df[full.df$record %in% exclusions.1, ]
ebsco.remove <- ebsco.remove[, "articleID"]
full.df <- full.df[!(full.df$articleID %in% ebsco.remove),]

rm(ebsco.remove)
```

## Exclusions of social work publications

Exclusions were then made any of the three criteria   

+ Journals that did not have clear evidence of a peer review process.  This was based on a review of the journals mission, scope, and whether an editorial board or group could be identified.  

+ Journal titles that did not have any clear publication history based on a broad-based Internet search and other databases (e.g., ProQuest).  These exclusions are assumed to be data entry problems at time of indexing.  

+ Journals that did not have at least one full year of indexed records (except journals that subsequently changed titles).  This exclusionary criterion was consistent with the working definition of a social work journal.  

```{r exclusion2, eval=TRUE}
exclusions.2 <- c(
    "Eugenics and Social Welfare Bulletin",
    "Eugenics & Social Welfare Bull.",
    "Issues in Social Work Education", 
    "Hospital Social Service",  #Confirmed nursing journal
    "International Conference of Social Work",
    "Journal of Social Work Process",
    "Maatskaplike Werk/Social Work",
    "Medical Social Work",
    "New York State Department of Social Welfare, Division Publication",
    "Pediatric Social Work", 
    "Pennsylvania Social Work",
    "Proceedings of the National Conference of Social Work","Social Service",    
    "Social Service Department, Presbyterian Hospital, New York",
    "Social Service Monographs",
    "School Social Work Quarterly", 
    "Social Work Today",
    "Social Work Technique",
    "Social Work Today", 
    "Social Work Yearbook",
    "Pennsylvania Social Work",
    "University of Chicago Social Service Monograph" 
    )

ebsco.remove <- full.df[full.df$record %in% exclusions.2, ]
ebsco.remove <- ebsco.remove[, "articleID"]
full.df <- full.df[!(full.df$articleID %in% ebsco.remove),]

rm(ebsco.remove, exclusions.2)
```

## Title merging

Some social work journals changed titles over their history.  For example, _Families in Society_ had three titles that preceded the current title:  _Social Casework_, _Journal of Social Casework_, and _The Family_.  The database that contributed the largest number of articles to the study database indexed the majority of journal titles under the journal's current name, and provided an additional data field for the historical name.  However, historical names from Social Science Abstracts and Social Work Abstracts were not merged with the current name.  Thus, the last step of the data cleaning procedures was to conduct these merges.  

```{r titleMerge, eval=TRUE}
full.df$record <- ifelse(full.df$record == 
    "The British Journal of Pyschiatric Social Work", # Old
        "British Journal of Social Work",           # New title
                        full.df$record)

full.df$record <- ifelse(full.df$record == 
        "Computers in Human Services",              # Old title
        "Journal of Technology in Human Services",  # New title
                        full.df$record)


full.df$record <- ifelse(full.df$record == 
        "The Family" | full.df$record ==                    # Old title
        "Journal of Social Casework" | full.df$record ==    # Old title
        "Social Casework",                                  # Old title
        "Families in Society",                              # New title
                        full.df$record)

full.df$record <- ifelse(full.df$record == 
        "Social Work in Education",                 # Old title
        "Children & Schools",                       # New title
                    full.df$record)

full.df$record <- ifelse(full.df$record == 
        "Journal of Independent Social Work" | full.df$record ==    # New title
        "Journal of Analytic Social Work",                          # Old title
        "Psychoanalytic Social Work",                               # Old title
                full.df$record)

full.df$record <- ifelse(full.df$record ==
        "Journal of Practice Teaching in Social Work and Practice", #Old title
        "Journal of Practice Teaching in Social Work and Health",   #New title
                    full.df$record)

# Journals were originally merged after suffixes were eliminated (after the
# colon).  Thus, Practice: Social Work in Action  (new title) was shortened to
# Practice (old title).  This section resolves two of these issues to be
# consistent in the comparison with Hodge.

practice.temp <- grep("^Practice$", full.df$record)
full.df$record[practice.temp] <- "Practice: Social Work in Action"

practice.temp <- grep("^Reflections$", full.df$record)
full.df$record[practice.temp] <- "Reflections: Narratives of Professional Helping"
```


# Post-processing summary 

## Revised database

This section provides a summary of the revised database after the initial processing with the `ebscoBWR` function and the subsequent data cleaning.  An additional code is included in this summary to indicate whether the journal title overlaps with the list of Hodge (indicated by *H*).  

```{r echo=FALSE, comment=NA}
unique.titles <- filter(full.df, attributes == "journal") %>%
    group_by(record) %>%
    summarise(N = n())

hodge <- readLines("/Users/beperron/Git/SocialWorkResearch/search_specifications/hodge.txt", n = -1)

unique.titles <- filter(full.df, attributes == "journal") %>%
    group_by(record) %>%
    mutate(Hodge.list = ifelse(record %in% hodge == TRUE, 1, 0)) %>%
    summarise(N = n(), 
              Hodge.list = max(Hodge.list)) %>%
    mutate(Hodge = ifelse(Hodge.list == 1, "Y", "N"))


n.so.yr <- filter(full.df, attributes == "journal" | attributes == "pubYear")

n.so <- filter(full.df, attributes == "journal") %>% mutate(title = record) %>% 
        select(-attributes, -record)

n.yr <- filter(full.df, attributes == "pubYear") %>% mutate(year = record ) %>% 
        select(-attributes, -record)

n.so.yr <- left_join(n.so, n.yr) %>%
    group_by(title) %>%
    summarise(first = min(year), last = max(year), n = n()) %>%
    arrange(title)

H <- unique.titles$Hodge

final.list <- cbind(n.so.yr, H)

print.data.frame(final.list, row.names=FALSE)
```




## Intersection, Differences and Union

This section provides further comparison of the journal titles with Hodge by reporting on the _intersection_, _differences_, and _union_ of the lists.
```{r}
hodge <- readLines("/Users/beperron/Git/SocialWorkResearch/search_specifications/hodge.txt", n = -1)

unique.swHistory <- filter(full.df, attributes == "journal") %>%
                distinct(record)

hodge.swHistory.intersect <- intersect(hodge, unique.swHistory$record)

hodge.swHistory.diff <- setdiff(hodge, unique.swHistory$record)
swHistory.hodge.diff <- setdiff(unique.swHistory$record, hodge)


hodge.swHistory.union <- union(hodge, unique.swHistory$record)
```


This list shows which journal titles are common to both lists.

```{r}
print(sort(hodge.swHistory.intersect))
```


This list shows the articles on Hodge but not the current study.  

```{r}
sink(file = "JournalsMissed.txt")
print(sort(hodge.swHistory.diff))
sink()
```
This list shows the articles on the current study but not on Hodge.

```{r}
print(sort(swHistory.hodge.diff))
```

This list shows the union of lists -- that is, a full list of all titles by combining a list.  It would be useful to break out the former titles of journals, as well as the years of existence.  This would establish a complete listing of all social work journals, both current and historical.  

```{r}
print(sort(hodge.swHistory.union))
```


## Descriptive summary of records

```{r comment=NA, echo=TRUE} 
number.of.articles <- length(which(full.df$attributes == "article"))
range.of.years <- range(full.df$record[full.df$attributes == "pubYear"])
```


```{r}
# Number of articles
cat(number.of.articles)
```

```{r}
# Range of years
cat(range.of.years)
```

```{r}
cor.data <- select(final.list, first, last, n) %>%
            mutate(no.years = as.numeric(last) - as.numeric(first)) 

ggplot(cor.data, aes(x = no.years, y = n)) + 
            geom_point(shape = 1) + 
            geom_smooth() + 
            xlab("Years") + 
            ylab("Number of articles")

cor(cor.data$no.years, cor.data$n)
```